---
title: "Network Analysis: Examining Fake News Echo Chambers Across Political Facebook Groups"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Installing Packages and Loading Libraries

```{r}
#install.packages("devtools")
library(devtools)

#install_github("pablobarbera/Rfacebook/Rfacebook")

#install.packages("Rfacebook") 
library(Rfacebook) # Facebook API
```

### Generating Token

```{r}
# token generated here: https://developers.facebook.com/tools/explorer 
token <- 'EAACEdEose0cBAA1FI8RxsAEXIQJHv9ghrm2bpNBJsNSSAtXQGYL18L5dTBKI2CpSMU0Pg0CJvmhsmX6wclGee7fVELTMQclPNdGOhRNDfY9LfaWI3AAXBhJ1ZBd9aV65415b99WQTtRTIZBJAADzDZCV3BvJk51TfUSR0ZCAxooZCa4mb6IZCv9fitZANe9KCcZD'
```

### Extracting posted links from selected political Facebook pages

```{r}
###Conservative/Right-Leaning Pages
TPPats <- getPage("teapartypatriots", token, n = 5000)
#saveRDS(TPPats, "TPPats.RDS")
#str(TPPats)
#TPPats$link
#TPPats$type

Trump <- getPage("DonaldTrump4President", token, n = 5000)

GOP <- getPage("GrandOldParty", token, n = 5000)

ConservDaily <- getPage("conservativedaily", token, n=5000)

EagleRising <- getPage("theEagleisRising", token, n=5000)

FreeDaily <- getPage("freedomdailynews", token, n=5000)

RightWing <- getPage("officialrightwingnews", token, n=5000)

###Liberal/Left-Leaning Pages
Hillary <- getPage("imwithhillary2016", token, n = 5000)

DemSocialists <- getPage("USdems", token, n = 5000)

CollegeDems<-getPage("collegedems", token, n = 5000)

Bernie <- getPage("PeopleForBernie", token, n = 5000)

BeingLiberal <- getPage("beingliberal.org", token, n = 5000)

OccupyDems <- getPage("OccupyDemocrats", token, n=5000)

Other98 <- getPage("theother98", token, n=5000)
```

### Shortening links to just domains

```{r}
#Function to extract link
domain <-function(x)strsplit(gsub("http.//|https://|www\\.", "",x), "/")

#Applying to Pages
TPPatsdom <- domain(TPPats$link)
TPPatsdom

Hillarydom <- domain(Hillary$link)

Trumpdom <- domain(Trump$link)

DemSocialistsdom <- domain(DemSocialists$link)

CollegeDemsdom <- domain(CollegeDems$link)

GOPdom <- domain(GOP$link)

Berniedom <- domain(Bernie$link)

BeingLiberaldom <- domain(BeingLiberal$link)

ConservDailydom <- domain(ConservDaily$link)

EagleRisingdom <- domain(EagleRising$link)

OccupyDemsdom<-domain(OccupyDems$link)

FreeDailydom <- domain(FreeDaily$link)

Other98dom <- domain(Other98$link)

RightWingdom <- domain(RightWing$link)

#
TPPats$domain<-c()
for (i in 1:length(TPPats$link)){
  TPPats$domain[i]<-TPPatsdom[[c(i,1)]]
}
print(TPPats$domain)


Hillary$domain<-c()
for (i in 1:length(Hillary$link)){
  Hillary$domain[i]<-Hillarydom[[c(i,1)]]
}

Trump$domain<-c()
for (i in 1:length(Trump$link)){
  Trump$domain[i]<-Trumpdom[[c(i,1)]]
}

DemSocialists$domain<-c()
for (i in 1:length(DemSocialists$link)){
  DemSocialists$domain[i]<-DemSocialistsdom[[c(i,1)]]
}

CollegeDems$domain<-c()
for (i in 1:length(CollegeDems$link)){
  CollegeDems$domain[i]<-CollegeDemsdom[[c(i,1)]]
}

GOP$domain<-c()
for (i in 1:length(GOP$link)){
  GOP$domain[i]<-GOPdom[[c(i,1)]]
}

Bernie$domain<-c()
for (i in 1:length(Bernie$link)){
  Bernie$domain[i]<-Berniedom[[c(i,1)]]
}

BeingLiberal$domain<-c()
for (i in 1:length(BeingLiberal$link)){
  BeingLiberal$domain[i]<-BeingLiberaldom[[c(i,1)]]
}

ConservDaily$domain<-c()
for (i in 1:length(ConservDaily$link)){
  ConservDaily$domain[i]<-ConservDailydom[[c(i,1)]]
}

EagleRising$domain<-c()
for (i in 1:length(EagleRising$link)){
  EagleRising$domain[i]<-EagleRisingdom[[c(i,1)]]
}

OccupyDems$domain<-c()
for (i in 1:length(OccupyDems$link)){
  OccupyDems$domain[i]<-OccupyDemsdom[[c(i,1)]]
}

FreeDaily$domain<-c()
for (i in 1:length(FreeDaily$link)){
  FreeDaily$domain[i]<-FreeDailydom[[c(i,1)]]
}

RightWing$domain<-c()
for (i in 1:length(RightWing$link)){
  RightWing$domain[i]<-RightWingdom[[c(i,1)]]
}

Other98$domain<-c()
for (i in 1:length(Other98$link)){
  Other98$domain[i]<-Other98dom[[c(i,1)]]
}
```

### Removing facebook.com and youtube.com domains and forming matrices

```{r}
a1 <- table(TPPats$domain)
a1 <- a1[!rownames(a1)=="facebook.com"]
a1 <- a1[!rownames(a1)=="youtube.com"]
#a1 <- a1[rownames(a1)%in%FakeNews$`100percentfedup.com`]
b1 <- as.matrix(a1)
colnames(b1) <- "TPPats"
sum(b1)

a3 <- table(Hillary$domain)
a3 <- a3[!rownames(a3)=="facebook.com"]
a3 <- a3[!rownames(a3)=="youtube.com"]
#a3 <- a3[rownames(a3)%in%FakeNews$`100percentfedup.com`]
b3 <- as.matrix(a3)
colnames(b3) <- "Hillary"

a4 <- table(Trump$domain)
a4 <- a4[!rownames(a4)=="facebook.com"]
#a4 <- a4[!rownames(a4)=="youtube.com"]
#a4 <- a4[rownames(a4)%in%FakeNews$`100percentfedup.com`]
b4 <- as.matrix(a4)
colnames(b4) <- "Trump"

a5 <- table(DemSocialists$domain)
a5 <- a5[!rownames(a5)=="facebook.com"]
a5 <- a5[!rownames(a5)=="youtube.com"]
#a5 <- a5[rownames(a5)%in%FakeNews$`100percentfedup.com`]
b5 <- as.matrix(a5)
colnames(b5) <- "DemSocialists"

a6 <- table(CollegeDems$domain)
a6 <- a6[!rownames(a6)=="facebook.com"]
a6 <- a6[!rownames(a6)=="youtube.com"]
#a6 <- a6[rownames(a6)%in%FakeNews$`100percentfedup.com`]
b6 <- as.matrix(a6)
colnames(b6) <- "CollegeDems"

a7 <- table(GOP$domain)
a7 <- a7[!rownames(a7)=="facebook.com"]
a7 <- a7[!rownames(a7)=="youtube.com"]
#a7 <- a7[rownames(a7)%in%FakeNews$`100percentfedup.com`]
b7 <- as.matrix(a7)
colnames(b7) <- "GOP"

a8 <- table(Bernie$domain)
a8 <- a8[!rownames(a8)=="facebook.com"]
a8 <- a8[!rownames(a8)=="youtube.com"]
#a8 <- a8[rownames(a8)%in%FakeNews$`100percentfedup.com`]
b8 <- as.matrix(a8)
colnames(b8) <- "Bernie"

a9 <- table(BeingLiberal$domain)
a9 <- a9[!rownames(a9)=="facebook.com"]
a8 <- a9[!rownames(a9)=="youtube.com"]
#a9 <- a9[rownames(a9)%in%FakeNews$`100percentfedup.com`]
b9 <- as.matrix(a9)
colnames(b9) <- "BeingLiberal"

a11 <- table(ConservDaily$domain)
a11 <- a11[!rownames(a11)=="facebook.com"]
a11 <- a11[!rownames(a11)=="youtube.com"]
#a11 <- a11[rownames(a11)%in%FakeNews$`100percentfedup.com`]
b11 <- as.matrix(a11)
colnames(b11) <- "ConservDaily"

a12 <- table(EagleRising$domain)
a12 <- a12[!rownames(a12)=="facebook.com"]
a12 <- a12[!rownames(a12)=="youtube.com"]
#a12 <- a12[rownames(a12)%in%FakeNews$`100percentfedup.com`]
b12 <- as.matrix(a12)
colnames(b12) <- "EagleRising"

a13 <- table(OccupyDems$domain)
a13 <- a13[!rownames(a13)=="facebook.com"]
a13 <- a13[!rownames(a13)=="youtube.com"]
#a13 <- a13[rownames(a13)%in%FakeNews$`100percentfedup.com`]
b13 <- as.matrix(a13)
colnames(b13) <- "OccupyDems"

a14 <- table(FreeDaily$domain)
a14 <- a14[!rownames(a14)=="facebook.com"]
a14 <- a14[!rownames(a14)=="youtube.com"]
#a14 <- a14[rownames(a14)%in%FakeNews$`100percentfedup.com`]
b14 <- as.matrix(a14)
colnames(b14) <- "FreeDaily"

a16 <- table(Other98$domain)
a16 <- a16[!rownames(a16)=="facebook.com"]
a16 <- a16[!rownames(a16)=="youtube.com"]
#a16 <- a16[rownames(a16)%in%FakeNews$`100percentfedup.com`]
b16 <- as.matrix(a16)
colnames(b16) <- "Other98"

a17 <- table(RightWing$domain)
a17 <- a17[!rownames(a17)=="facebook.com"]
a17 <- a17[!rownames(a17)=="youtube.com"]
#a17 <- a17[rownames(a17)%in%FakeNews$`100percentfedup.com`]
b17 <- as.matrix(a17)
colnames(b17) <- "RightWing"

```

### Removing non-fakenews domains

```{r}
a1 <- table(TPPats$domain)
a1 <- a1[!rownames(a1)=="facebook.com"]
a1 <- a1[!rownames(a1)=="youtube.com"]
a1 <- a1[rownames(a1)%in%FakeNews$`X__1`]
b1 <- as.matrix(a1)
colnames(b1) <- "TPPats"
sum(b1)

a3 <- table(Hillary$domain)
a3 <- a3[!rownames(a3)=="facebook.com"]
a3 <- a3[!rownames(a3)=="youtube.com"]
a3 <- a3[rownames(a3)%in%FakeNews$`X__1`]
b3 <- as.matrix(a3)
colnames(b3) <- "Hillary"

a4 <- table(Trump$domain)
a4 <- a4[!rownames(a4)=="facebook.com"]
a4 <- a4[!rownames(a4)=="youtube.com"]
a4 <- a4[rownames(a4)%in%FakeNews$`X__1`]
b4 <- as.matrix(a4)
colnames(b4) <- "Trump"

a5 <- table(DemSocialists$domain)
a5 <- a5[!rownames(a5)=="facebook.com"]
a5 <- a5[!rownames(a5)=="youtube.com"]
a5 <- a5[rownames(a5)%in%FakeNews$`X__1`]
b5 <- as.matrix(a5)
colnames(b5) <- "DemSocialists"

a6 <- table(CollegeDems$domain)
a6 <- a6[!rownames(a6)=="facebook.com"]
a6 <- a6[!rownames(a6)=="youtube.com"]
a6 <- a6[rownames(a6)%in%FakeNews$`X__1`]
b6 <- as.matrix(a6)
colnames(b6) <- "CollegeDems"

a7 <- table(GOP$domain)
a7 <- a7[!rownames(a7)=="facebook.com"]
a7 <- a7[!rownames(a7)=="youtube.com"]
a7 <- a7[rownames(a7)%in%FakeNews$`X__1`]
b7 <- as.matrix(a7)
colnames(b7) <- "GOP"

a8 <- table(Bernie$domain)
a8 <- a8[!rownames(a8)=="facebook.com"]
a8 <- a8[!rownames(a8)=="youtube.com"]
a8 <- a8[rownames(a8)%in%FakeNews$`X__1`]
b8 <- as.matrix(a8)
colnames(b8) <- "Bernie"

a9 <- table(BeingLiberal$domain)
a9 <- a9[!rownames(a9)=="facebook.com"]
a8 <- a9[!rownames(a9)=="youtube.com"]
a9 <- a9[rownames(a9)%in%FakeNews$`X__1`]
b9 <- as.matrix(a9)
colnames(b9) <- "BeingLiberal"

a11 <- table(ConservDaily$domain)
a11 <- a11[!rownames(a11)=="facebook.com"]
a11 <- a11[!rownames(a11)=="youtube.com"]
a11 <- a11[rownames(a11)%in%FakeNews$`X__1`]
b11 <- as.matrix(a11)
colnames(b11) <- "ConservDaily"

a12 <- table(EagleRising$domain)
a12 <- a12[!rownames(a12)=="facebook.com"]
a12 <- a12[!rownames(a12)=="youtube.com"]
a12 <- a12[rownames(a12)%in%FakeNews$`X__1`]
b12 <- as.matrix(a12)
colnames(b12) <- "EagleRising"

a13 <- table(OccupyDems$domain)
a13 <- a13[!rownames(a13)=="facebook.com"]
a13 <- a13[!rownames(a13)=="youtube.com"]
a13 <- a13[rownames(a13)%in%FakeNews$`X__1`]
b13 <- as.matrix(a13)
colnames(b13) <- "OccupyDems"

a14 <- table(FreeDaily$domain)
a14 <- a14[!rownames(a14)=="facebook.com"]
a14 <- a14[!rownames(a14)=="youtube.com"]
a14 <- a14[rownames(a14)%in%FakeNews$`X__1`]
b14 <- as.matrix(a14)
colnames(b14) <- "FreeDaily"

a16 <- table(Other98$domain)
a16 <- a16[!rownames(a16)=="facebook.com"]
a16 <- a16[!rownames(a16)=="youtube.com"]
a16 <- a16[rownames(a16)%in%FakeNews$`X__1`]
b16 <- as.matrix(a16)
colnames(b16) <- "Other98"

a17 <- table(RightWing$domain)
a17 <- a17[!rownames(a17)=="facebook.com"]
a17 <- a17[!rownames(a17)=="youtube.com"]
a17 <- a17[rownames(a17)%in%FakeNews$`X__1`]
b17 <- as.matrix(a17)
colnames(b17) <- "RightWing"
```

### Make each page it's own network

```{r}
library(igraph)
c1 <- graph_from_incidence_matrix(b1, weighted = T)
c1

c3 <- graph_from_incidence_matrix(b3, weighted=T)
c3
c4 <- graph_from_incidence_matrix(b4, weighted=T)
c5 <- graph_from_incidence_matrix(b5, weighted=T)
c6 <- graph_from_incidence_matrix(b6, weighted=T)
c7 <- graph_from_incidence_matrix(b7, weighted=T)
c8 <- graph_from_incidence_matrix(b8, weighted=T)
c9 <- graph_from_incidence_matrix(b9, weighted=T)

c11 <- graph_from_incidence_matrix(b11, weighted=T)
c12 <- graph_from_incidence_matrix(b12, weighted=T)
c13 <- graph_from_incidence_matrix(b13, weighted=T)
c14 <- graph_from_incidence_matrix(b14, weighted=T)
c16 <- graph_from_incidence_matrix(b16, weighted=T)
c17 <- graph_from_incidence_matrix(b17, weighted=T)
```

### Combine all
# d1 is now a network of both pages and linked sites

```{r}
d1 <- union(c1, c3, c4, c5, c6, c7, c8, c9, c11, c12, c13, c14, c16, c17)

plot(d1, vertex.size = V(d1)$deg, vertex.label = NA, edge.width = E(d1)$weight)
E(d1)
```

### Mark whether each  node is a page or linked site

```{r}
V(d1)$type <- bipartite.mapping(d1)$type
V(d1)$type
```

### Make network of pages (Pages as Nodes, Shared Links as Edges)

```{r}
e1 <- bipartite.projection(d1)$proj2
E(e1)$weight

vertex_attr( e1, 'deg' ) <- degree( e1 , mode = 'total' )

vertex_attr(e1, 'domain') <- sapply(strsplit(V(e1)$name , "\\." ), 'tail', 1 )
```

### Examining Network

```{r}
table(e1$weight)
E(e1)$weight
```

### Labeling Nodes (Pages) as Either Republican (Conservative) or Democrat (Liberal)

```{r}
V(e1)$party <- c("republican", "democrat", "republican", "democrat", "democrat", "republican", "democrat", "democrat", "republican", "republican", "democrat", "republican", "democrat", "republican")

# Marking Conservative Nodes as Red and Liberal Nodes as Blue
E(e1)$color <- c("red", "red", "purple", "purple", "purple", "purple", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "blue", "red", "purple", "red", "red", "purple", "blue", "blue", "blue", "blue", "red", "red", "red", "red", "blue", "red")

# Visualizing Network
plot(e1, layout=layout_with_kk, vertex.size = V(e1)$deg, vertex.label = (V(e1)$domain), vertex.label.color="black", edge.arrow.mode = '-', vertex.color=c( "red", "blue")[1+(V(e1)$party=="democrat")], edge.width=(E(e1)$weight)*1.5, edge.color=E(e1)$color) 
```

### Density and centralization measures for the network

```{r}
graph.density(e1)

centralization.betweenness(e1)$centralization

centralization.degree(e1)$centralization

centralization.closeness(e1)$centralization
```

### Clustering Network (Political Groups) based on Shared Links (Weighted Edges)

```{r}
#clustering
com <- edge.betweenness.community(e1)
V(e1)$memb <- com$membership
modularity(com)
com <- edge.betweenness.community(e1)

plot(com, e1)

plot( com1, e3, vertex.size = V(e3)$deg, vertex.label = (V(e3)$domain), vertex.label.color="black", edge.arrow.mode = '-', vertex.color=c( "red", "blue")[1+(V(e3)$party=="democrat")], edge.width=log(E(e3)$weight)*2, edge.color=E(e3)$color) 
```

